Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-25 parallel.tex
working on unigrid scaling section

% Brian to work on this, scaling data from Harkness
Figure~\ref{fig.uniscale} shows \enzo\ weak scaling for a sequence of
scaled non-AMR Lyman alpha forest calculations on NICS Kraken XT4 and
ORNL Jaguar XT4 using decomposition with $128^3$ root grid tiles.  The
grid resolution is of constant physical size.  Scaling up to 98,304
cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
version of \enzo\.  Excellent scaling in unigrid mode is seen as long
as each core is given enough work -- somewhere around $128^3$
cells/core seems to be about right.

>>>>

It is advantageous to use \enzo\ in its ``unigrid'' (i.e.,
non-adaptive mesh) mode for a variety of problems, including
non-gravitating turbulence
\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04}, the Lyman-alpha forest
\citep{2005MNRAS.361...70J,2009MNRAS.399.1934P}, or feedback of
metal-enriched gas from galaxies into the intergalactic medium
\citep{2004ApJ...601L.115N,2011ApJ...731....6S}.  Achieving good
scaling of the code in unigrid mode is relatively straightforward --
upon initialization, unigrid simulations are decomposed such that each
MPI process has a roughly equal subvolume (and thus number of grid
cells), meaning that work is evenly distributed.  Communication
patterns for both the gravity solve (which uses a fast Fourier
transform) and the fluid solves (which transfer boundary information
between subvolumes) are predictable and straightforward, and
rebuilding of the grid hierarchy does not take place, removing a
substantial global operation and a great deal of communication.

Figure~\ref{fig.uniscale} shows \enzo\ weak scaling results for a
sequence of scaled unigrid Lyman alpha forest calculations. These
calculations include dark matter dynamics, hydrodynamics using the
piecewise parabolic method, six-species non-equilibrium chemistry and
radiative cooling, and a uniform metagalactic ultraviolet background.
In this sequence of test calculations, we perform a weak scaling test
on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
supercomputers\footnote{these simulations were performed in 2008,
prior to conversion of both machines to the current-generation
systems}.  In this test, each MPI task was given a $128^3$ root grid
tile (i.e., $128^3$ grid cells containing baryon quantities) and,
initially, approximately $128^3$ dark matter particles.  The number of
grid cells was constant throughout the calculation; the number of dark
matter particles varies as they are moved from subvolume to subvolume
as structure evolves.  The grid resolution was kept at a constant
comoving size of $\simeq 40$~kpc/h, and as the core count was
increased, so was the simulation volume.  On each machine, a compute
node contained a single AMD Opteron quad-core chip (2.1 Ghz on Jaguar;
2.3 Ghz on Kraken) with 2 GB/memory per core (8 GB/total per node).
Both machines used the SeaStar2 interconnect.  In the scaling study,
calculations were run with 1, 2 or 4 MPI tasks per node.  The figure
shows cell updates per second per MPI process; perfect scaling would
be a horizontal line.

As can be seen in Figure~\ref{fig.uniscale}, the unigrid weak scaling
performance of the code is extremely good for this problem, with only
a 20\% decrease in cell updates/second/task as the code is scaled from
8 to 4,096 MPI tasks, and a 40\% decrease in performance overall going
from 8 to 13,824 (or $24^3$) MPI tasks.  We speculate that this
decrease is likely to be partially due to global MPI communications
used to, e.g., calculate the overall timestep of the simulation, and
also likely due to load imbalances due to increasing cosmological
power (and thus an increasingly uneven distribution of dark matter
particles between MPI tasks at late times) as the simulation volume
grows.  We also observe that a systematic difference in speed can be
seen between the two machines, which can be attributed primarily to
the slightly faster CPUs on Kraken (2.3 Ghz, vs. 2.1 Ghz on Jaguar).
The difference in speed when using different numbers of MPI tasks per
node can be attributed primarily due to differences in competing usage
of shared cache on the quad-core chips used on this machine.

Broadly, excellent scaling in \enzo's unigrid
mode is seen for a variety of problems as long as each compute core is
given an adequate amount of work to do.  For cosmological simulations, this seems to be
roughly $128^3$ cells per core.  If fewer cells per core are used, the
CPU is essentially data-starved, and poor scaling is observed due to
computing units being idle while waiting for information to be
communicated from other processes (for, e.g., boundary information or
gravity solves).  Substantially larger cell counts per core  would in
principle help scaling, but are typically impractical on
most machines due to memory limits.

As a final point, we 
up to 98,304
cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
(MPI + OpenMP)
version of \enzo\.  Excellent scaling in unigrid mode is seen as long
as each core is given enough work -- somewhere around $128^3$
cells/core seems to be about right.
